Fusion Moves for Correlation Clustering

### 2014-11-13 paper/egpaper_for_review.tex
update 5.1
update 6/7

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage{bm}% ändert \boldsymbol
\def\arraystretch{0.8}
\renewcommand{\tabcolsep}{2pt}
\newcommand{\thickline}{2pt}
\newcommand{\scatterplotpath}{./scatterplots/}

@@

\subsection{Improvements for the Multicut Algorithm}\label{sec:tunedMC} 
When using the  public available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal~\cite{kappes_2013_arxiv}, we realize that their implementation has some limitations on large problems.
This results in a very slow separation and make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.
In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.

@@

\begin{table}[t]
   \tiny
   \centering
   \caption{Evaluation of Variation of Information (VI) for datasets with available ground-truth.}
   \label{tab:eval}
   \begin{tabular}{lllllll}
      \toprule
         Dataset          &  image-seg   & knott-3d-150 & knott-3d-300   & knott-3d-450 & knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &   $ 4.9633$  &    $ 2.9936$ &    $ 4.4986$   &      --      &    --        &  --\\ 
         HC               &   $ 2.5967$  &    $ 1.5477$ &    $ 2.3513$   &    $ 2.9155$ &    --        & $       2.8395$\\
         HC-CGC           &   $ 2.5164$  &    $ 0.9052$ &    $ 1.7636$   &    $ 2.2256$ &    --        & $       1.7603$ \\
         CGC              &   $ 2.5247$  &    $ 0.9267$ &    $ 1.8822$   &    $ 2.3104$ &    --        & $       6.8908$ \\
         KL               &   $ 2.6432$  &    $ 2.0648$ &    $ 4.1318$   &    $ 4.9270$ &    --        & $       7.1057$\\
         FUSION           &$\bf 2.1406$  &    $ 2.8787$ &    $ 4.0744$   &    $ 4.6616$ &    --        & $       6.5366$ \\
         MC-R             &   $ 2.5471$  &    $ 0.9178$ &    $ 1.6369$   &    $ 2.8710$ &    --        & $       6.5058$\\   
         MC-I             &   $ 2.5367$  & $\bf 0.9063$ & $\bf 1.6352$   & $\bf 2.0037$ &    --        & $       4.3319$\\  
         FusionCC-HC-MC   &   $ 2.5319$  &    $ 0.9629$ &    $ 1.6516$   &    $ 2.0801$ &    --        & $\bf    1.3347$ \\  
         FusionCC-HC-CGC  &$\bf 2.4961$  &    $ 0.9679$ &    $ 1.7673$   &    $ 2.3809$ &    --        & $       2.1347$ \\  
         FusionCC-WS-MC   &   $ 2.5340$  &    $ 0.9629$ &    $ 1.6742$   &    $ 2.0739$ &    --        & $\bf    1.3334$ \\  
         FusionCC-WS-CGC  &   $ 2.5192$  &    $ 1.0585$ &    $ 2.1344$   &    $ 2.7487$ &    --        & $       3.3514$ \\      
      \bottomrule
   \end{tabular}
\end{table}

\begin{table}[t]
   \tiny
   \centering
   \caption{Evaluation of Rand Index (RI) for datasets with available ground-truth.}
   \label{tab:eval2}
   \begin{tabular}{lllllll}
      \toprule
         Dataset          &  image-seg   & knott-3d-150 & knott-3d-300   & knott-3d-450 & knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &   $ 0.7438$  &    $ 0.7851$ &    $  0.8792$  &      --      &    --        &  --\\ 
         HC               &   $ 0.7560$  &    $ 0.8139$ &    $  0.8084$  &    $ 0.7610$ &    --        & $ 0.9651$\\
         HC-CGC           &   $ 0.7724$  &    $ 0.9226$ &    $  0.8713$  &    $ 0.8433$ &    --        & $ 0.9861$ \\
         CGC              &   $ 0.7590$  &    $ 0.9206$ &    $  0.8666$  &    $ 0.8341$ &    --        & $ 0.6024$ \\
         KL               &   $ 0.6400$  &    $ 0.8085$ &    $  0.6858$  &    $ 0.6409$ &    --        & $ 0.5849$\\
         FUSION           &   $ 0.5480$  &    $ 0.2849$ &    $  0.1420$  &    $ 0.0998$ &    --        & $ 0.0345$ \\
         MC-R             &   $ 0.7822$  &    $ 0.9232$ & $\bf  0.8849$  &    $ 0.6713$ &    --        & $ 0.0432$ \\   
         MC-I             &   $ 0.7821$  & $\bf 0.9236$ & $\bf  0.8849$  & $\bf 0.8670$ &    --        & $ 0.5461$\\  
         FusionCC-HC-MC   &   $ 0.7801$  &    $ 0.9042$ &    $  0.8824$  &    $ 0.8573$ &    --        & $\bf 0.9906$ \\  
         FusionCC-HC-CGC  &   $ 0.7780$  &    $ 0.9031$ &    $  0.8763$  &    $ 0.8470$ &    --        & $ 0.9775$ \\  
         FusionCC-WS-MC   &$\bf 0.7825$  &    $ 0.9042$ &    $  0.8802$  &    $ 0.8582$ &    --        & $ 0.8895$ \\  
         FusionCC-WS-CGC  &   $ 0.7750$  &    $ 0.8951$ &    $  0.8596$  &    $ 0.8394$ &    --        & $\bf 0.9906$ \\      
      \bottomrule
   \end{tabular}
\end{table}

@@

\section{Future Work}\label{sec:future}
In future work we would like to investigate in much larger problem instances.
Within this challenging task we can also enforce additionally must-not-link constraints in eq.~\ref{eq:fusion_move_a}.
If those build a closed surface, the problem decompose into two independent problems.
This is also useful for interactive correlation clustering, by enforcing
updates only in local regions by appropriate proposals.
Overall, a more specific selection of proposals, maybe conditioned on the current solution,
should lead better proposals and to faster progress of CCFusion.
%
Furthermore, we can extend our approach to higher-order correlation clustering~\cite{Kim-2011,kappes_2013_arxiv}
by using higher order subproblems.

@@

We have presented a fast and scalable 
approximate solver for correlation 
clustering, named Correlation Clustering Fusion (CCFusion).
The best solution is iteratively improved 
by a fusion with proposal solutions.
The fusion move itself is formulated as correlation
clustering on a smaller graph fewer edges and nodes
and can therefore be solved much faster than the original problem.

>>>>

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage{bm}% ändert \boldsymbol
\def\arraystretch{0.8}
\renewcommand{\tabcolsep}{2pt}
\newcommand{\thickline}{2pt}
\newcommand{\scatterplotpath}{./scatterplots/}
\newcommand{\rd}{\color{red}}

@@

\subsection{Improvements for the Multicut Algorithm}\label{sec:tunedMC} 
When using the  public available implementation in OpenGM\footref{opengm-master}
of Kappes~\etal~\cite{kappes_2013_arxiv}, we realize that their implementation has some limitations on large problems.
This results in a very slow separation and we make the following modifications.
Firstly, we use \emph{index-min-heap}~\cite{Sedgewick-2011} within the shortest path search by the Dykstra algorithm - this gives a speed up for MC-R.
Secondly, we follow~\cite{kroeger_2012_eccv} and search for shortest paths and add those only if they are non-chordal instead of searching for the shortest non-chordal path
during the separation procedure. 
In~\cite{kroeger_2012_eccv}. this has been used for MC-I only. 
For MC-R this search procedure is not sufficient and need to be followed by a search for a shortest non-chordal paths.
%
Fig.~\ref{fig:at:knott-450b} show the improvements with our modifications compared to the implementation in OpenGM for the knott-3d-450 dataset.
This procedure is one magnitude faster, but might cause a few extra outer iterations.

@@

\begin{table}[t]
   \tiny
   \centering
   \caption{Evaluation by Variation of Information (VOI) Rand Index (RI) for datasets with available ground-truth.}
   \label{tab:eval}
   \begin{tabular}{lllllll}
      \toprule
         VOI              &  image-seg   & knott-3d-150 & knott-3d-300   & knott-3d-450 & knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &\rd$ 4.9633$  &\rd $ 2.9936$ &\rd $ 4.4986$   &      --      &    --        &  --\\ 
         HC               &   $ 2.5967$  &    $ 1.5477$ &    $ 2.3513$   &    $ 2.9155$ &    --        & $       2.8395$\\
         HC-CGC           &   $ 2.5164$  &    $ 0.9052$ &    $ 1.7636$   &    $ 2.2256$ &    --        & $       1.7603$ \\
         CGC              &   $ 2.5247$  &    $ 0.9267$ &    $ 1.8822$   &    $ 2.3104$ &    --        & $\rd    6.8908$ \\
         KL               &   $ 2.6432$  &\rd $ 2.0648$ &\rd $ 4.1318$   &\rd $ 4.9270$ &    --        & $\rd    7.1057$\\
         FUSION           &$\bf 2.1406$  &\rd $ 2.8787$ &\rd $ 4.0744$   &\rd $ 4.6616$ &    --        & $\rd    6.5366$ \\
         MC-R             &   $ 2.5471$  &    $ 0.9178$ &    $ 1.6369$   &    $ 2.8710$ &    --        & $\rd    6.5058$\\   
         MC-I             &   $ 2.5367$  & $\bf 0.9063$ & $\bf 1.6352$   & $\bf 2.0037$ &    --        & $\rd    4.3319$\\  
         FusionCC-HC-MC   &   $ 2.5319$  &    $ 0.9629$ &    $ 1.6516$   &    $ 2.0801$ &    --        & $\bf    1.3347$ \\  
         FusionCC-HC-CGC  &$\bf 2.4961$  &    $ 0.9679$ &    $ 1.7673$   &    $ 2.3809$ &    --        & $       2.1347$ \\  
         FusionCC-WS-MC   &   $ 2.5340$  &    $ 0.9629$ &    $ 1.6742$   &    $ 2.0739$ &    --        & $\bf    1.3334$ \\  
         FusionCC-WS-CGC  &   $ 2.5192$  &    $ 1.0585$ &    $ 2.1344$   &    $ 2.7487$ &    --        & $       3.3514$ \\      
      \bottomrule
   \end{tabular}

   \begin{tabular}{lllllll}
      \toprule
          RI              &  image-seg   & knott-3d-150 & knott-3d-300   & knott-3d-450 & knott-3d-550 &3d-seg\\
      \midrule 
         PIVOT-BOEM       &   $ 0.7438$  &\rd $ 0.7851$ &    $  0.8792$  &      --      &    --        &  --\\ 
         HC               &   $ 0.7560$  &\rd $ 0.8139$ &\rd $  0.8084$  &\rd $ 0.7610$ &    --        &   $ 0.9651$\\
         HC-CGC           &   $ 0.7724$  &    $ 0.9226$ &    $  0.8713$  &    $ 0.8433$ &    --        &   $ 0.9861$ \\
         CGC              &   $ 0.7590$  &    $ 0.9206$ &    $  0.8666$  &    $ 0.8341$ &    --        &\rd$ 0.6024$ \\
         KL               &\rd$ 0.6400$  &\rd $ 0.8085$ &\rd $  0.6858$  &\rd $ 0.6409$ &    --        &\rd$ 0.5849$\\
         FUSION           &\rd$ 0.5480$  &\rd $ 0.2849$ &\rd $  0.1420$  &\rd $ 0.0998$ &    --        &\rd$ 0.0345$ \\
         MC-R             &   $ 0.7822$  &    $ 0.9232$ & $\bf  0.8849$  &\rd $ 0.6713$ &    --        &\rd$ 0.0432$ \\   
         MC-I             &   $ 0.7821$  & $\bf 0.9236$ & $\bf  0.8849$  & $\bf 0.8670$ &    --        &\rd$ 0.5461$\\  
         FusionCC-HC-MC   &   $ 0.7801$  &    $ 0.9042$ &    $  0.8824$  &    $ 0.8573$ &    --        &$\bf 0.9906$ \\  
         FusionCC-HC-CGC  &   $ 0.7780$  &    $ 0.9031$ &    $  0.8763$  &    $ 0.8470$ &    --        &   $ 0.9775$ \\  
         FusionCC-WS-MC   &$\bf 0.7825$  &    $ 0.9042$ &    $  0.8802$  &    $ 0.8582$ &    --        &   $ 0.8895$ \\  
         FusionCC-WS-CGC  &   $ 0.7750$  &    $ 0.8951$ &    $  0.8596$  &    $ 0.8394$ &    --        &   $\bf 0.9906$ \\      
      \bottomrule
   \end{tabular}
\end{table}

@@

\section{Future Work}\label{sec:future}
In future work we would like to investigate in much larger problem instances
and interactive correlation clustering, by enforcing
updates only in local regions by appropriate proposals.
%
Within this challenging task we can also enforce additionally must-not-link constraints in eq.~\ref{eq:fusion_move_a}.
If those build a closed surface, the problem decompose into two independent problems.
%
Overall, a more specific selection of proposals, maybe conditioned on the current solution,
should lead better proposals and to faster progress of CCFusion.
%
Furthermore, we can extend our approach to higher-order correlation clustering~\cite{Kim-2011,kappes_2013_arxiv}
by using higher order subproblems.

@@

We have presented a fast and scalable 
approximate solver for correlation 
clustering, named Correlation Clustering Fusion (CCFusion).
It is orthogonal to previous research, \ie it can be combined with and
correlation clustering solver.
%
The best solution is iteratively improved 
by a fusion with proposal solutions.
The fusion move itself is formulated as correlation
clustering on a smaller graph fewer edges and nodes
and can therefore be solved much faster than the original problem.
